---
title: "Heterogeneity of CD8 T cells (Kang 2017) with Modules"
author: "Nathan Lawlor"
date: "5/29/2018"
output: html_document
---

## Load packages and data
```{r load_packages, echo=TRUE, message=FALSE}
rm(list = ls())
library(Seurat)
library(iasva)
library(SummarizedExperiment)
library(irlba)
library(Rtsne)
library(RColorBrewer)
library(pheatmap)
library(corrplot)
library(tmod)
library(tagcloud)
library(limma)
library(VennDiagram)
#setwd("/Users/lawlon/Documents/CZI/Seurat_Public_Data/immune_alignment_expression_matrices/")
name <- "Kang.2017.all.cells"
pct_cutt <- 1
ctrl.data <- readRDS("/Users/lawlon/Documents/CZI/Seurat_Public_Data/immune_alignment_expression_matrices/immune_control_expression_matrix.Rds")
dim(ctrl.data)
stim.data <- readRDS("/Users/lawlon/Documents/CZI/Seurat_Public_Data/immune_alignment_expression_matrices/immune_stimulated_expression_matrix.Rds")
dim(stim.data)
# normalize function
normalize <- function (counts) 
{
  normfactor <- colSums(counts)
  return(t(t(counts)/normfactor)*median(normfactor))
}
color.vec <- c("#e41a1c", "#377eb8", "#4daf4a", "#984ea3", "#ff7f00", "black", "#a65628", "#f781bf", "#999999")
```

## Filter genes based on over-dispersion
```{r filter_genes, echo=TRUE, message=FALSE, eval=TRUE}
# load in seurat annotations
imm.comb <- readRDS("/Users/lawlon/Documents/CZI/Seurat_Public_Data/immune_alignment_expression_matrices/Seurat/Kang.2017.PBMC.stimulated.Seurat.immune.combined.Rds")
imm.meta <- imm.comb@meta.data
imm.meta <- cbind(imm.meta, imm.comb@ident)
colnames(imm.meta)[ncol(imm.meta)] <- "celltype"
dim(imm.meta)
all.meta <- imm.meta
imm.meta <- imm.meta[imm.meta$celltype == "CD8 T" & imm.meta$stim == "CTRL",]

# Set up control object
ctrl <- CreateSeuratObject(raw.data = ctrl.data[, rownames(imm.meta)], project = "IMMUNE_CTRL", min.cells = 5)
ctrl@meta.data$stim <- "CTRL"
ctrl <- FilterCells(ctrl, subset.names = "nGene", low.thresholds = 500, high.thresholds = Inf)
ctrl <- NormalizeData(ctrl)
ctrl <- ScaleData(ctrl, display.progress = F)

# Gene selection for input to CCA
ctrl <- FindVariableGenes(ctrl, do.plot = F)
g.1 <- head(rownames(ctrl@hvg.info), 2000)
genes.use <- g.1

# combine data
ctrl.data.sel <- ctrl.data[genes.use, rownames(ctrl@meta.data)]
all.counts <- ctrl.data.sel
dim(all.counts)
# overdispersed genes
#saveRDS(all.counts, file = "immune_control_stimulated_filtered_counts.Rds")

```

## Genes Detected (CTRL cells)
```{r filter_data, echo=TRUE, message=FALSE, fig.width=7, fig.height=7}
# calculate lib size
#pdf(file = paste(name, cell_interest, "IASVA.genes.detected.hist.pdf", sep = "."))
Num_Detected_Genes <- colSums(all.counts > 0)
hist(Num_Detected_Genes, col = "dodgerblue4", breaks = 20, ylab = "Number of Cells")
summ <- summary(Num_Detected_Genes)
legend("topright", legend = paste(names(summ), summ))
abline(v = 500, col = "red", lty = 2)
legend("right", legend = paste("< 500 genes = ", length(which(Num_Detected_Genes < 500)), " Cells\n",
                                  "> 500 genes = ", length(which(Num_Detected_Genes > 500)), " Cells", sep = ""))
#dev.off()
```

## Known Factors (Ctrl Cells)
Here, the only known factors are stimulation, celltype, and library size
```{r known_factors, echo=TRUE, message=FALSE}
# make vector of condition
Condition <- factor(imm.meta$stim)
CellType <- factor(imm.meta$celltype)
# normalize counts
counts.filt <- normalize(all.counts)
Geo_Lib <- colSums(log(counts.filt+1))
# categorize lib size
Geo_Lib_Cat <- cut(Num_Detected_Genes, breaks = 4)
summary(Geo_Lib)
barplot(Geo_Lib, xlab="Cell", las=2, ylab = "Geometric Library Size")               
lcounts <- log(counts.filt + 1)
lcounts <- as.matrix(lcounts)
# PC1 and Geometric library size correlation
pc1 = irlba(lcounts - rowMeans(lcounts), 1)$v[,1] ## partial SVD
cor(Geo_Lib, pc1)
```

## T-SNE prior to IASVA analysis
```{r run_pre_tsne, echo=TRUE, message=FALSE, fig.width=7, fig.height=7}
tsne.res <- Rtsne(unique(t(lcounts)), dims = 2)
#saveRDS(tsne.res, file = paste(name, "tsne.preIASVA.Rds", sep = "."))

# by celltype
par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
plot(tsne.res$Y[,1:2], main="Pre IA-SVA", xlab="tSNE Dim1",
     ylab="tSNE Dim2", pch=20, col=color.vec[CellType], 
     bg=color.vec[CellType], cex = 0.5)
par(xpd=TRUE)
legend(50,20, levels(CellType), border="white", fill=color.vec, bty="n", title = "Cell Type")
# by condition
par(mar=c(5, 4, 4, 2), xpd=FALSE)
plot(tsne.res$Y[,1:2], main="Pre IA-SVA", xlab="tSNE Dim1",
     ylab="tSNE Dim2", pch=20, col=color.vec[Condition], 
     bg=color.vec[Condition], oma=c(4,4,6,12), cex = 0.5)
legend("topright", levels(Condition), border="white", fill=color.vec, bty="n", title = "Condition")
# by genes detect
plot(tsne.res$Y[,1:2], main="Pre IA-SVA", xlab="tSNE Dim1",
     ylab="tSNE Dim2", pch=20, col=color.vec[Geo_Lib_Cat], 
     bg=color.vec[Geo_Lib_Cat], oma=c(4,4,6,12), cex = 0.5)
legend("topright", levels(Geo_Lib_Cat), border="white", fill=color.vec, bty="n", title = "Genes Detected")

```

## Run IASVA using num.sv = 10 (Ctrl cells)
```{r run_iasva, echo=TRUE, message=FALSE}
# run iasva
mod <- model.matrix(~Geo_Lib)
summ_exp <- SummarizedExperiment(assays = as.matrix(counts.filt))
iasva.res<- fast_iasva(summ_exp, mod[,-1, drop = F], verbose=FALSE, pct.cutoff = pct_cutt, num.sv = 10)
# load results
iasva.sv <- iasva.res$sv
#saveRDS(iasva.res, file = paste(name, "IASVA", pct_cutt, "output.Rds", sep = "."))
markers_pct1 <- iasva::find_markers(Y = summ_exp, iasva.sv = iasva.sv)
# revised find markers function
source("/Users/lawlon/iasva_markers_update/R/find_markers.R")
all_markers <- .GlobalEnv$find_markers(Y = summ_exp, iasva.sv = iasva.sv)
# all_markers <- readRDS("/Users/lawlon/Documents/CZI/Seurat_Public_Data/immune_alignment_expression_matrices/IASVA/All_Cells/Kang.2017.all.cells.IASVA.by.SV.markers.Rds")

#saveRDS(markers_pct1, file = paste(name, "IASVA", pct_cutt, "markers.Rds", sep = "."))
#tsne on markers
set.seed(100)
if (ncol(markers_pct1) > 100) {
  perplex <- 30
} else {
  perplex <- round(ncol(markers_pct1)/4)
}
tsne.res.iasva <- Rtsne(unique(t(log(markers_pct1+1))), dims = 2, perplexity = perplex)
# saveRDS(tsne.res.iasva, file = paste(name, cell_interest, stim_cond, "IASVA", pct_cutt, "tsne.markers.Rds", sep = "."))
```

## Visualize SVs (Ctrl cells)
```{r plot_svs, echo=TRUE, message=FALSE, fig.width=8, fig.height=8}
ncol(iasva.sv)
# if (ncol(iasva.sv) > 5) {
#   pairs(iasva.sv[,1:5], main="IA-SVA", pch=20, col=color.vec[CellType],
#       bg=color.vec[CellType], cex=0.5, oma=c(4,4,6,12))
#   legend("right", levels(CellType), border="white", fill=color.vec, bty="n")
#   pairs(iasva.sv[,1:5], main="IA-SVA", pch=20, col=color.vec[Condition],
#       bg=color.vec[Condition], cex=0.5, oma=c(4,4,6,12))
#   legend("right", levels(Condition), border="white", fill=color.vec, bty="n")
#   pairs(iasva.sv[,1:5], main="IA-SVA", pch=20, col=color.vec[Geo_Lib_Cat],
#       bg=color.vec[Geo_Lib_Cat], cex=0.5, oma=c(4,4,6,12))
#   legend("right", levels(Geo_Lib_Cat), border="white", fill=color.vec, bty="n")
# } else {
  par(mar=c(5.1, 4.1, 4.1, 8.2), xpd=TRUE)
  pairs(iasva.sv, main="IA-SVA", pch=20, col=color.vec[CellType],
      bg=color.vec[CellType], cex=0.5, oma=c(4,4,6,12))
  legend(0.85,0.5, levels(CellType), border="white", fill=color.vec, bty="n")
  pairs(iasva.sv, main="IA-SVA", pch=20, col=color.vec[Condition],
      bg=color.vec[Condition], cex=0.5, oma=c(4,4,6,12))
  legend("right", levels(Condition), border="white", fill=color.vec, bty="n")
  pairs(iasva.sv, main="IA-SVA", pch=20, col=color.vec[Geo_Lib_Cat],
      bg=color.vec[Geo_Lib_Cat], cex=0.5, oma=c(4,4,6,12))
  legend("right", levels(Geo_Lib_Cat), border="white", fill=color.vec, bty="n", title = "Genes Detected")
  
# }

```

## Correlate SVS (Ctrl cells)
```{r corr_svs, echo=TRUE, message=FALSE, fig.width=7, fig.height=7}
iasva_vars <- cbind(iasva.sv, Geo_Lib)
corrplot(cor(iasva_vars))

```

## Modules enriched in each SV (Ctrl cells)
```{r modules, echo=TRUE, message=FALSE, fig.width=9, fig.height=9}
# iterative analysis of SVs an modules
all_sv_res <- list()
par(mfrow = c(4,2), mar = c(4,2,2,2))
for (j in 1:ncol(iasva.sv)) {
    sv_genes <- all_markers[[j]]
    design <- model.matrix(~iasva.sv[,j])
    fit <- eBayes(lmFit(counts.filt, design))
    tt <- topTable(fit, coef=2, number=Inf, genelist = rownames(counts.filt))
    
    # module analysis
    fg <- tt[sv_genes,]
    fg.ord <- fg[order(fg$logFC, decreasing = T),]
    res <- tmodHGtest(fg = fg.ord$ID, bg = tt$ID)
    if (is.null(res)) {}
    else {
      res.ord <- res[order(res$E, decreasing = T),]
    # append results
    all_sv_res[[j]] <- res.ord
    names(all_sv_res)[j] <- names(all_markers)[j]
    if (nrow(res.ord) > 1) {
      # make plots
    bp1 <- barplot(res.ord$E, main = paste(names(all_markers)[j], " Module Enrichment", sep = ""), xlab = "", horiz = T,
                  cex.main = 2, cex.lab = 2)
    #barplot(res.ord$E, xlab = "Module Enrichment", main = paste(names(all_markers)[j], " Genes", sep = ""), horiz = T)
    par(xpd=TRUE)
    text(x = max(res.ord$E)/2, y = bp1[,1], labels = paste(res.ord$Title, " (", res$b, "/", res$B, ")", sep = ""), cex = 1, font = 2)
    }
    }
}

```

## Plot All Markers (Ctrl cells)
```{r plot_markers, echo=TRUE, message=FALSE, fig.width=7, fig.height=7, eval = TRUE}
anno.col <- data.frame(Condition = Condition, Genes_Detect = Geo_Lib_Cat, CellType = CellType)
rownames(anno.col) <- colnames(markers_pct1)
# how many markers are there
dim(markers_pct1)
pheatmap(log(markers_pct1+1), show_colnames = FALSE, show_rownames = TRUE,
         clustering_method = "ward.D2", fontsize_row = 10,
         annotation_col = anno.col)

```

## TSNE using IASVA Marker genes (by genes detected) (Ctrl cells)
```{r tsne_gene_detect, echo=TRUE, fig.width=7, fig.height=7}
# load results
par(mfrow = c(1,1))
par(mar=c(5.1, 4.1, 4.1, 8.2), xpd=TRUE)
plot(tsne.res.iasva$Y[,1:2], main="IA-SVA", xlab="tSNE Dim1",
     ylab="tSNE Dim2", pch=20, col=color.vec[CellType], 
     bg=color.vec[CellType], oma=c(4,4,6,12), cex = 0.5)
legend(40, 20, levels(CellType), border="white", fill=color.vec, bty="n")
par(mar=c(5, 4, 4, 2), xpd=FALSE)
plot(tsne.res.iasva$Y[,1:2], main="IA-SVA", xlab="tSNE Dim1",
     ylab="tSNE Dim2", pch=20, col=color.vec[Geo_Lib_Cat], 
     bg=color.vec[Geo_Lib_Cat], oma=c(4,4,6,12), cex = 0.5)
legend("topright", levels(Geo_Lib_Cat), border="white", fill=color.vec, bty="n", title = "Genes Detected")
plot(tsne.res.iasva$Y[,1:2], main="IA-SVA", xlab="tSNE Dim1",
     ylab="tSNE Dim2", pch=20, col=color.vec[Condition], 
     bg=color.vec[Condition], oma=c(4,4,6,12), cex = 0.5)
legend("topright", levels(Condition), border="white", fill=color.vec, bty="n")

```


## ANALYSIS with Stimulated Cells
```{r filter_genes_stim, echo=TRUE, message=FALSE, eval=TRUE}
# load in seurat annotations
imm.comb <- readRDS("/Users/lawlon/Documents/CZI/Seurat_Public_Data/immune_alignment_expression_matrices/Seurat/Kang.2017.PBMC.stimulated.Seurat.immune.combined.Rds")
imm.meta <- imm.comb@meta.data
imm.meta <- cbind(imm.meta, imm.comb@ident)
colnames(imm.meta)[ncol(imm.meta)] <- "celltype"
dim(imm.meta)
all.meta <- imm.meta
imm.meta <- imm.meta[imm.meta$celltype == "CD8 T" & imm.meta$stim == "STIM",]

# Set up control object
stim <- CreateSeuratObject(raw.data = stim.data[, rownames(imm.meta)], project = "IMMUNE_STIM", min.cells = 5)
stim@meta.data$stim <- "STIM"
stim <- FilterCells(stim, subset.names = "nGene", low.thresholds = 500, high.thresholds = Inf)
stim <- NormalizeData(stim)
stim <- ScaleData(stim, display.progress = F)

# Gene selection for input to CCA
stim <- FindVariableGenes(stim, do.plot = F)
g.2 <- head(rownames(stim@hvg.info), 2000)
genes.use <- g.2

# combine data
stim.data.sel <- stim.data[genes.use, rownames(stim@meta.data)]
all.counts <- stim.data.sel
dim(all.counts)
# overdispersed genes
#saveRDS(all.counts, file = "immune_control_stimulated_filtered_counts.Rds")

```

## Genes Detected (STIM cells)
```{r filter_data_stim, echo=TRUE, message=FALSE, fig.width=7, fig.height=7}
# calculate lib size
#pdf(file = paste(name, cell_interest, "IASVA.genes.detected.hist.pdf", sep = "."))
Num_Detected_Genes <- colSums(all.counts > 0)
hist(Num_Detected_Genes, col = "dodgerblue4", breaks = 20, ylab = "Number of Cells")
summ <- summary(Num_Detected_Genes)
legend("topright", legend = paste(names(summ), summ))
abline(v = 500, col = "red", lty = 2)
legend("right", legend = paste("< 500 genes = ", length(which(Num_Detected_Genes < 500)), " Cells\n",
                                  "> 500 genes = ", length(which(Num_Detected_Genes > 500)), " Cells", sep = ""))
#dev.off()
```

## Known Factors (STIM Cells)
Here, the only known factors are stimulation, celltype, and library size
```{r known_factors_stim, echo=TRUE, message=FALSE}
# make vector of condition
Condition <- factor(imm.meta$stim)
CellType <- factor(imm.meta$celltype)
# normalize counts
counts.filt <- normalize(all.counts)
Geo_Lib <- colSums(log(counts.filt+1))
# categorize lib size
Geo_Lib_Cat <- cut(Num_Detected_Genes, breaks = 4)
summary(Geo_Lib)
barplot(Geo_Lib, xlab="Cell", las=2, ylab = "Geometric Library Size")               
lcounts <- log(counts.filt + 1)
lcounts <- as.matrix(lcounts)
# PC1 and Geometric library size correlation
pc1 = irlba(lcounts - rowMeans(lcounts), 1)$v[,1] ## partial SVD
cor(Geo_Lib, pc1)
```

## T-SNE prior to IASVA analysis (STIM cells)
```{r run_pre_tsne_stim, echo=TRUE, message=FALSE, fig.width=7, fig.height=7}
tsne.res <- Rtsne(unique(t(lcounts)), dims = 2)
#saveRDS(tsne.res, file = paste(name, "tsne.preIASVA.Rds", sep = "."))

# by celltype
par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
plot(tsne.res$Y[,1:2], main="Pre IA-SVA", xlab="tSNE Dim1",
     ylab="tSNE Dim2", pch=20, col=color.vec[CellType], 
     bg=color.vec[CellType], cex = 0.5)
par(xpd=TRUE)
legend(50,20, levels(CellType), border="white", fill=color.vec, bty="n", title = "Cell Type")
# by condition
par(mar=c(5, 4, 4, 2), xpd=FALSE)
plot(tsne.res$Y[,1:2], main="Pre IA-SVA", xlab="tSNE Dim1",
     ylab="tSNE Dim2", pch=20, col=color.vec[Condition], 
     bg=color.vec[Condition], oma=c(4,4,6,12), cex = 0.5)
legend("topright", levels(Condition), border="white", fill=color.vec, bty="n", title = "Condition")
# by genes detect
plot(tsne.res$Y[,1:2], main="Pre IA-SVA", xlab="tSNE Dim1",
     ylab="tSNE Dim2", pch=20, col=color.vec[Geo_Lib_Cat], 
     bg=color.vec[Geo_Lib_Cat], oma=c(4,4,6,12), cex = 0.5)
legend("topright", levels(Geo_Lib_Cat), border="white", fill=color.vec, bty="n", title = "Genes Detected")

```

## Run IASVA using num.sv = 10 (Stim cells)
```{r run_iasva_stim, echo=TRUE, message=FALSE}
# run iasva
mod <- model.matrix(~Geo_Lib)
summ_exp <- SummarizedExperiment(assays = as.matrix(counts.filt))
iasva.res<- fast_iasva(summ_exp, mod[,-1, drop = F], verbose=FALSE, pct.cutoff = pct_cutt, num.sv = 10)
# load results
iasva.sv <- iasva.res$sv
#saveRDS(iasva.res, file = paste(name, "IASVA", pct_cutt, "output.Rds", sep = "."))
markers_pct1 <- iasva::find_markers(Y = summ_exp, iasva.sv = iasva.sv)
# revised find markers function
source("/Users/lawlon/iasva_markers_update/R/find_markers.R")
all_markers <- .GlobalEnv$find_markers(Y = summ_exp, iasva.sv = iasva.sv)
# all_markers <- readRDS("/Users/lawlon/Documents/CZI/Seurat_Public_Data/immune_alignment_expression_matrices/IASVA/All_Cells/Kang.2017.all.cells.IASVA.by.SV.markers.Rds")

#saveRDS(markers_pct1, file = paste(name, "IASVA", pct_cutt, "markers.Rds", sep = "."))
#tsne on markers
set.seed(100)
if (ncol(markers_pct1) > 100) {
  perplex <- 30
} else {
  perplex <- round(ncol(markers_pct1)/4)
}
tsne.res.iasva <- Rtsne(unique(t(log(markers_pct1+1))), dims = 2, perplexity = perplex)
# saveRDS(tsne.res.iasva, file = paste(name, cell_interest, stim_cond, "IASVA", pct_cutt, "tsne.markers.Rds", sep = "."))
```

## Visualize SVs (Stim cells)
```{r plot_svs_stim, echo=TRUE, message=FALSE, fig.width=8, fig.height=8}
ncol(iasva.sv)
# if (ncol(iasva.sv) > 5) {
#   pairs(iasva.sv[,1:5], main="IA-SVA", pch=20, col=color.vec[CellType],
#       bg=color.vec[CellType], cex=0.5, oma=c(4,4,6,12))
#   legend("right", levels(CellType), border="white", fill=color.vec, bty="n")
#   pairs(iasva.sv[,1:5], main="IA-SVA", pch=20, col=color.vec[Condition],
#       bg=color.vec[Condition], cex=0.5, oma=c(4,4,6,12))
#   legend("right", levels(Condition), border="white", fill=color.vec, bty="n")
#   pairs(iasva.sv[,1:5], main="IA-SVA", pch=20, col=color.vec[Geo_Lib_Cat],
#       bg=color.vec[Geo_Lib_Cat], cex=0.5, oma=c(4,4,6,12))
#   legend("right", levels(Geo_Lib_Cat), border="white", fill=color.vec, bty="n")
# } else {
  par(mar=c(5.1, 4.1, 4.1, 8.2), xpd=TRUE)
  pairs(iasva.sv, main="IA-SVA", pch=20, col=color.vec[CellType],
      bg=color.vec[CellType], cex=0.5, oma=c(4,4,6,12))
  legend(0.85,0.5, levels(CellType), border="white", fill=color.vec, bty="n")
  pairs(iasva.sv, main="IA-SVA", pch=20, col=color.vec[Condition],
      bg=color.vec[Condition], cex=0.5, oma=c(4,4,6,12))
  legend("right", levels(Condition), border="white", fill=color.vec, bty="n")
  pairs(iasva.sv, main="IA-SVA", pch=20, col=color.vec[Geo_Lib_Cat],
      bg=color.vec[Geo_Lib_Cat], cex=0.5, oma=c(4,4,6,12))
  legend("right", levels(Geo_Lib_Cat), border="white", fill=color.vec, bty="n", title = "Genes Detected")
  
# }

```

## Correlate SVS (Stim cells)
```{r corr_svs_stim, echo=TRUE, message=FALSE, fig.width=7, fig.height=7}
iasva_vars <- cbind(iasva.sv, Geo_Lib)
corrplot(cor(iasva_vars))

```

## Modules enriched in each SV (Stim cells)
```{r modules_stim, echo=TRUE, message=FALSE, fig.width=9, fig.height=9}
# iterative analysis of SVs an modules
all_sv_res <- list()
par(mfrow = c(4,2), mar = c(4,2,2,2))
for (j in 1:ncol(iasva.sv)) {
    sv_genes <- all_markers[[j]]
    design <- model.matrix(~iasva.sv[,j])
    fit <- eBayes(lmFit(counts.filt, design))
    tt <- topTable(fit, coef=2, number=Inf, genelist = rownames(counts.filt))
    
    # module analysis
    fg <- tt[sv_genes,]
    if (nrow(fg) > 1) {
      fg.ord <- fg[order(fg$logFC, decreasing = T),]
    } else {
      fg.ord <- fg
    }
    
    res <- tmodHGtest(fg = fg.ord$ID, bg = tt$ID)
    if (is.null(res)) {}
    else {
      res.ord <- res[order(res$E, decreasing = T),]
    # append results
    all_sv_res[[j]] <- res.ord
    names(all_sv_res)[j] <- names(all_markers)[j]
    if (nrow(res.ord) > 1) {
      # make plots
    bp1 <- barplot(res.ord$E, main = paste(names(all_markers)[j], " Module Enrichment", sep = ""), xlab = "", horiz = T,
                  cex.main = 2, cex.lab = 2)
    #barplot(res.ord$E, xlab = "Module Enrichment", main = paste(names(all_markers)[j], " Genes", sep = ""), horiz = T)
    par(xpd=TRUE)
    text(x = max(res.ord$E)/2, y = bp1[,1], labels = paste(res.ord$Title, " (", res$b, "/", res$B, ")", sep = ""), cex = 1, font = 2)
    }
    }
    
}

```

## Plot All Markers (Stim cells)
```{r plot_markers_stim, echo=TRUE, message=FALSE, fig.width=7, fig.height=7, eval = TRUE}
anno.col <- data.frame(Condition = Condition, Genes_Detect = Geo_Lib_Cat, CellType = CellType)
rownames(anno.col) <- colnames(markers_pct1)
# how many markers are there
dim(markers_pct1)
pheatmap(log(markers_pct1+1), show_colnames = FALSE, show_rownames = TRUE,
         clustering_method = "ward.D2", fontsize_row = 10,
         annotation_col = anno.col)

```

## TSNE using IASVA Marker genes (by genes detected) (Ctrl cells)
```{r tsne_gene_detect_stim, echo=TRUE, fig.width=7, fig.height=7}
# load results
par(mfrow = c(1,1))
par(mar=c(5.1, 4.1, 4.1, 8.2), xpd=TRUE)
plot(tsne.res.iasva$Y[,1:2], main="IA-SVA", xlab="tSNE Dim1",
     ylab="tSNE Dim2", pch=20, col=color.vec[CellType], 
     bg=color.vec[CellType], oma=c(4,4,6,12), cex = 0.5)
legend(40, 20, levels(CellType), border="white", fill=color.vec, bty="n")
par(mar=c(5, 4, 4, 2), xpd=FALSE)
plot(tsne.res.iasva$Y[,1:2], main="IA-SVA", xlab="tSNE Dim1",
     ylab="tSNE Dim2", pch=20, col=color.vec[Geo_Lib_Cat], 
     bg=color.vec[Geo_Lib_Cat], oma=c(4,4,6,12), cex = 0.5)
legend("topright", levels(Geo_Lib_Cat), border="white", fill=color.vec, bty="n", title = "Genes Detected")
plot(tsne.res.iasva$Y[,1:2], main="IA-SVA", xlab="tSNE Dim1",
     ylab="tSNE Dim2", pch=20, col=color.vec[Condition], 
     bg=color.vec[Condition], oma=c(4,4,6,12), cex = 0.5)
legend("topright", levels(Condition), border="white", fill=color.vec, bty="n")

```


## Session Information
```{r session_info, echo=TRUE, message=FALSE}
sessionInfo()
```

